Simulation-WTCCC-2

### 2020-03-25 analysis/simulation-WTCCC.Rmd
fix results display bug

$$ \tilde{X} = G\alpha + \xi$$
For each gene, we sample $L$ eQTLs for this gene, for eQTL $k$, we have
$$ \alpha_k \sim N(0,\sigma_\alpha^2) $$
$$ \xi \sim N(0,1) $$
Then we have the heritability of the gene:

@@

@@

Thus in order to get desired $PVE_{SNP}$ and $PVE_expr$, we set $\sigma_\theta^2$ and $\sigma_\gamma^2$ based on the following formula:

@@

```{r}
simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191111/"
show_res <- function(simutag){
  load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c , 1))
  row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")
  res <- readLines(paste0(simudir, simutag, "_VC/output/", simutag, ".log.txt"))
  outdf1$est <- as.numeric(strsplit(res[24], "  ")[[1]][2:4])
  outdf1$est.se <- as.numeric(strsplit(res[25], "  ")[[1]][2:4])
  print(outdf1)
  outdf2 <- data.frame("est"= c(as.numeric(strsplit(res[20], "  ")[[1]][2:3]), as.numeric(strsplit(res[22], "=")[[1]][2])))
  row.names(outdf2) <- c("PVE.expr","PVE.snp","PVE.total")
  outdf2$est.se <- c(as.numeric(strsplit(res[21], "  ")[[1]][2:3]), as.numeric(strsplit(res[23], "=")[[1]][2]))
  print(outdf2)
}
```

>>>>

$$ X' = G\alpha$$
For each gene, we sample $L$ eQTLs for this gene, for eQTL $k$, we have
$$ \alpha_k \sim N(0,\sigma_\alpha^2) $$
$$ \xi \sim N(0,1) $$
Then we have the heritability of the gene:

@@

We simulate phenotype following

$$Y = G\alpha\gamma + G\theta + \epsilon$$.

For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$). $\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$. To choose proper $\sigma_\gamma$ and $\sigma_\theta$. We used the following formula:

@@

Thus in order to get desired $PVE_{SNP}$ and $PVE_{expr}$, we set $\sigma_\theta^2$ and $\sigma_\gamma^2$ based on the following formula:

@@

```{r}
simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191111/"
load("/home/simingz/causalTWAS/WTCCC/bd.toy.RData")
Gvar <- mean(apply(X,2,var))
show_res <- function(simutag){
  load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c/Gvar , 1))
  row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")
  res <- readLines(paste0(simudir, simutag, "_VC/output/", simutag, ".log.txt"))
  outdf1$est <- as.numeric(strsplit(res[24], "  ")[[1]][2:4])
  outdf1$est.se <- as.numeric(strsplit(res[25], "  ")[[1]][2:4])
  print(outdf1)
  outdf2 <- data.frame("est"= c(as.numeric(strsplit(res[20], "  ")[[1]][2:3]), as.numeric(strsplit(res[22], "=")[[1]][2])))
  row.names(outdf2) <- c("PVE.expr","PVE.snp","PVE.total")
  outdf2$est.se <- c(as.numeric(strsplit(res[21], "  ")[[1]][2:3]), as.numeric(strsplit(res[23], "=")[[1]][2]))
  print(outdf2)
}
```

### 2020-03-25 analysis/simulation-WTCCC2.Rmd
fix results display bug

Simulation of phenotype: use the same weights as used in simulate expression using the `'cad.RData` dataset.

@@

## Simulate cis-expression 
We simulate cis-expression based on the following model:

@@

## Simulate quantitative phenotype
We simulate phenotype followin

@@

```{r}
simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191127/"
show_res <- function(simutag){
  load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c , 1))
  row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")
  res <- readLines(paste0(simudir, simutag, "_VC/output/", simutag, ".log.txt"))
  outdf1$est <- as.numeric(strsplit(res[24], "  ")[[1]][2:4])
  outdf1$est.se <- as.numeric(strsplit(res[25], "  ")[[1]][2:4])
  print(outdf1)
  outdf2 <- data.frame("est"= c(as.numeric(strsplit(res[20], "  ")[[1]][2:3]), as.numeric(strsplit(res[22], "=")[[1]][2])))
  row.names(outdf2) <- c("PVE.expr","PVE.snp","PVE.total")
  outdf2$est.se <- c(as.numeric(strsplit(res[21], "  ")[[1]][2:3]), as.numeric(strsplit(res[23], "=")[[1]][2]))
  print(outdf2)
}
```

@@

>>>>

Simulation of phenotype: use the same weights as used in simulate expression using the `'cad.RData` dataset, simulate phenotype according to:

$$Y = G\alpha\gamma + G\theta + \epsilon$$. For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$). $\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$.

@@

## Simulate cis-expression 
We simulate cis-expression based on the following model, using 500 samples from cad in WTCCC:

@@

## Simulate quantitative phenotype
We simulate phenotype using bd data from WTCCC following
$$Y = G\alpha\gamma + G\theta + \epsilon$$. For $\alpha$, we used the same as used in simulating expression (Thus, not $\tilde{X}$). $\gamma \sim N(0,\sigma_\gamma^2), \theta \sim N(0, \sigma_\theta^2), \epsilon \sim N (0,1)$.

@@

```{r}
simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191127/"
load("/home/simingz/causalTWAS/WTCCC/bd.toy.RData")
Gvar <- mean(apply(X,2,var))
show_res <- function(simutag){
  load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c/Gvar , 1))
  row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")
  res <- readLines(paste0(simudir, simutag, "_VC/output/", simutag, ".log.txt"))
  outdf1$est <- as.numeric(strsplit(res[24], "  ")[[1]][2:4])
  outdf1$est.se <- as.numeric(strsplit(res[25], "  ")[[1]][2:4])
  print(outdf1)
  outdf2 <- data.frame("est"= c(as.numeric(strsplit(res[20], "  ")[[1]][2:3]), as.numeric(strsplit(res[22], "=")[[1]][2])))
  row.names(outdf2) <- c("PVE.expr","PVE.snp","PVE.total")
  outdf2$est.se <- c(as.numeric(strsplit(res[21], "  ")[[1]][2:3]), as.numeric(strsplit(res[23], "=")[[1]][2]))
  print(outdf2)
}
```

@@

## Discussion
In this analysis, we see the PVE estimation for expression is much under estimated, this is because many of the gene expression training model has low $R^2$, especially for low heritability ones. For a polygenic model, because we are estimating the total herability explained by expression, the ones with low $R^2$ will affect this estimation. We are moving to model with a sparse prior for expression.
